---
title: "Regression analysis of MTCARS"
author: "Stanislav Prikhodko"
date: "Friday, May 22, 2015"
output: html_document
---



# Overview

Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and *miles per gallon* (MPG) (outcome). They are particularly interested in the following two questions:

- "Is an automatic or manual transmission better for MPG"

- "Quantify the MPG difference between automatic and manual transmissions"

# Exploratory Data Analysis

```{r}
data(mtcars)
head(mtcars)

boxplot(mtcars$mpg ~ mtcars$am, 
        col = "blue",
        xlab="Transmission (0 = Automatic, 1 = Manual)", 
        ylab="MPG",main="MPG over Transmission")
```

# Model Selection and Interpretation of coefficients
```{r}
model <- lm(mpg ~ am, data = mtcars)
summary(model)
```
Since *p-value* is very small (**0.000285**) the variable *am* appears to be a good predictor for *mpg*. We are 95% confident that there is a difference of mean of mpg among cars with automatic and manual transmission. 

# Residuals Plot

```{r}
par(mfrow = c(2, 2))
plot(model)
```

# Quantify the MPG difference 

In otder to quantify differenct of *MPG* between automatic and manual transmissions, let's calculate mean of two subsets
```{r}
automatic <- subset(mtcars, am==0)
manual <- subset(mtcars, am==1)
mpg_a <- mean(automatic$mpg)
mpg_m <- mean(manual$mpg)
difference <- mpg_m - mpg_a
```
The difference between *mpg* of manual and automatic cars in our sample is `r difference`. Let's perform t-test to estimate the difference in the whole population:
```{r}
t.test(mpg ~ am, data = mtcars)
```

Since *p-value* (0.001374) is less than 0.05, we reject the null hypothesis in favor of the alternative hypothesis. In ither words, we are 95 cconfident that thr true mean of mpg of automatic and manual cars (in this order) is between 3.209684 and 11.280194.

# Apendix I.

```{r}
summary(mtcars)
```
