---
title: "Regression analysis of MTCARS"
author: "Stanislav Prikhodko"
date: "Friday, May 22, 2015"
output: pdf_document
---



# Overview

Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and *miles per gallon* (MPG) (outcome). They are particularly interested in the following two questions:

- "Is an automatic or manual transmission better for MPG"

- "Quantify the MPG difference between automatic and manual transmissions"

# Exploratory Data Analysis

```{r}
data(mtcars)
head(mtcars)

boxplot(mtcars$mpg ~ mtcars$am, 
        xlab="Transmission (0 = Automatic, 1 = Manual)", 
        ylab="MPG",main="MPG over Transmission Type")
```

# Model Selection
```{r}
model <- lm(mpg ~ am, data = mtcars)
summary(model)
```
Since *p-value* is very small (**0.000285**) the variable *am* appears to be a good predictor for *mpg*.

# Residual Plot

```{r}
par(mfrow = c(2, 2))
plot(model)
```

# Apendix I.

```{r}
summary(mtcars)
```
