---
title: "Simulate CLT"
author: "Stanislav Prikhodko"
date: "Sunday, May 10, 2015"
output: html_document
---

```{r, echo=FALSE,warning=FALSE,include=FALSE}
library(ggplot2)
library(grid)

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

#http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
```

# Part 1. Sampling Exponential Distribution

## Overview

In this project I will investigate the exponential distribution and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with *rexp(n, lambda)* where lambda is the rate parameter. I will investigate the distribution of averages of 40 exponentials.

## Project  

The mean of exponential distribution is $\frac{1}{\lambda}$ and the standard deviation is also $\frac{1}{\lambda}$. I will use lambda = 0.2 for all of the simulations. In this study I will investigate the distribution of averages of 40 exponentials.

## Theoretical Mean of the distribution

Creating 1000 values from the exponential distribution
```{r}
lambda <- 0.2
exp <- rexp(n=1000, rate = lambda)
exp_mean <- mean(exp)
theoretical_mean <- (1/lambda)
```

The sample mean is **```r exp_mean```**. Theoretical mean is  **```r theoretical_mean```** (calculated as 1/lambda). 

## 

Let's plot a histogram to show the sample mean and compare it to the theoretical mean of the distribution.
```{r}
ggplot() + aes(exp)+ 
  geom_histogram( binwidth=1, colour="black", fill="white") + 
  geom_vline(aes(xintercept=exp_mean), color="red", linetype="dashed", size=1) 

```

### Theoretical Variance of the Sample

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r}
sample_size <- 40 
mns <- rep(0,1000)
for (i in 1:1000) {
  mns[i] <- mean(exp[sample(1:1000, sample_size, replace=FALSE)])
}

sample_var <- var(mns)
theoretical_var <- (1/lambda)^2 / sample_size

```

## Distribution is approximately normal

```{r, warning=FALSE} 
 
p1 <- 
  ggplot() + aes(mns)+ 
  geom_histogram(aes(y=..density..), binwidth=0.1, colour="black", fill="white") + 
  geom_vline(aes(xintercept=mean(mns, na.rm=T)), color="red", linetype="dashed", size=1) +  
  geom_density(alpha=.2)
print(p1)
```
```{r,include=FALSE}
p2 <- 
  ggplot() + aes(mns)+ 
  geom_histogram( binwidth=0.1, colour="black", fill="white") + 
  geom_vline(aes(xintercept=mean(mns, na.rm=T)), color="red", linetype="dashed", size=1)  

multiplot(p1,p2,cols=2)


 
```
 
 
