---
title: "Simulate CLT"
author: "Stanislav Prikhodko"
date: "Sunday, May 10, 2015"
output: html_document
---

```{r, echo=FALSE,warning=FALSE}
library(ggplot2)
library(grid)

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

#http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
```

# Part 1. Sampling Exponential Distribution

## Overview

In this project I will investigate the exponential distribution and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with *rexp(n, lambda)* where lambda is the rate parameter. I will investigate the distribution of averages of 40 exponentials.

## Project  

The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. I will use lambda = 0.2 for all of the simulations. In this study I will investigate the distribution of averages of 40 exponentials.

### Theoretical mean of the distribution

Creating 1000 values from the exponential distribution
```{r}
exp <- rexp(n=1000, rate = 0.2)
exp_mean <- mean(exp)
```

The sample mean is ```r exp_mean```. Let's plot a histogram to show the sample mean and compare it to the theoretical mean of the distribution.
```{r}
ggplot() + aes(exp)+ 
  geom_histogram( binwidth=1, colour="black", fill="white") + 
  geom_vline(aes(xintercept=exp_mean), color="red", linetype="dashed", size=1) 

```

#
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r}
exp[sample(1:5, 5, replace=FALSE)]
```

#
3. Show that the distribution is approximately normal.


```{r, warning=FALSE} 
 
mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(runif(40)))

p1 <- 
  ggplot() + aes(mns)+ 
  geom_histogram(aes(y=..density..), binwidth=0.01, colour="black", fill="white") + 
  geom_vline(aes(xintercept=mean(mns, na.rm=T)), color="red", linetype="dashed", size=1) +  
  geom_density(alpha=.2)

p2 <- 
  ggplot() + aes(mns)+ 
  geom_histogram( binwidth=0.01, colour="black", fill="white") + 
  geom_vline(aes(xintercept=mean(mns, na.rm=T)), color="red", linetype="dashed", size=1)  

multiplot(p1,p2,cols=2)


exp <- rexp(n=1000, rate = 0.2)
exp_mean <- mean(exp)

ggplot() + aes(exp)+ 
  geom_histogram( binwidth=1, colour="black", fill="white") + 
  geom_vline(aes(xintercept=exp_mean), color="red", linetype="dashed", size=1) 

#hist(mns)
```
 


## Sample Project Report Structure

Of course, there are multiple ways one could structure a report to address the requirements above.  However, the more clearly you pose and answer each question, the easier it will be for reviewers to clearly identify and evaluate your work. 

A sample set of headings that could be used to guide the creation of your report might be:

Simulations: Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.
Sample Mean versus Theoretical Mean: Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.
Sample Variance versus Theoretical Variance: Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.
Distribution: Via figures and text, explain how one can tell the distribution is approximately normal.

