---
title: "Practical Machine Learning: Writeup"
output: 
  html_document:
    keep_md: true
---

I will use the package **Caret** for predictive machine learning study.
```{r libraries,warning=FALSE,cache=FALSE}
library(caret)
library(corrplot)
library(kernlab)
library(randomForest)
```

## Loading and cleaning data

The following data will be used for analysis:
```{r,cache=TRUE}
if (!file.exists("pml-training.csv")) {
  urlTraining <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  urlTesting <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
  
  download.file(urlTraining , destfile = "pml-training.csv")
  download.file(urlTesting , destfile = "pml-testing.csv")
}
```

Loading raw data from the file:
```{r,echo=TRUE}
rawTraining <- read.csv("pml-training.csv", na.strings= c("NA",""," "))
rawTesting <- read.csv("pml-testing.csv", na.strings= c("NA",""," "))
```

Removing columns with missing data in it, or if it is not numeric:
```{r,echo=TRUE}
Training <- rawTraining[, colSums(is.na(rawTraining)) == 0] 
Testing <- rawTesting[, colSums(is.na(rawTraining)) == 0]
NumCols <- sapply(Training, is.numeric)
NumCols["classe"] <- TRUE
NumCols["raw_timestamp_part_1"] <- FALSE
NumCols["raw_timestamp_part_2"] <- FALSE
NumCols["X"] <- FALSE
NumCols["num_window"] <- FALSE
Training <- Training[, NumCols] 
Testing <- Testing[, NumCols]
# View(Training[1:20,]) 
```

## Slicing Data

I will use 80% of data as a training sample and the rest of data will be used for cross valudation:
```{r,echo=TRUE} 
inTrain <- createDataPartition(y = Training$classe, p = 0.8, list = FALSE)
trainingSet <- Training[inTrain, ]
crossVal <- Training[-inTrain, ]
correlMatrix <- cor(trainingSet[, -length(trainingSet)])
corrplot(correlMatrix, order="FPC", method="ellipse", type="upper", tl.col="black", tl.cex=0.5)
```

Then I run *random forest method* to create a model and build the confusion matrix for it: 
```{r,echo=TRUE}
model <- randomForest(classe ~ ., data = trainingSet)
predictCrossVal <- predict(model, crossVal)
confusionMatrix(crossVal$classe, predictCrossVal)
```

Then, predict the classes of the test set:
```{r, echo=FALSE}
predictTest <- predict(model, Testing)
print(predictTest)
```

The following code will run the model and save results into the file:
```{r,echo=FALSE}
answers = rep("A", 20)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
