---
title: "Impact of Severe Weather Events on Public Health and Economy in the US"
author: "Stas Prihodco"
date: "Wednesday, April 15, 2015"
output: html_document
---

# Synopsis 
  
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern. This study aims to identify which types of events are most harmful with respect to population health and which types of events have the greatest economic consequences.

# Data Processing

For this study I use [National Weather Service Storm Data,https://a.a.a].
Downloading data (**stormData.csv.bz2**) woud require *R.utils* package:
```{r, echo=TRUE}
library(R.utils)
if (!("repdata-data-StormData.csv.bz2" %in% dir(getwd()))) { 
 a<-1 #  download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", destfile = "repdata-data-StormData.csv.bz2")
}
if (!("repdata-data-StormData.csv" %in% dir(getwd()))) { 
 a<-1 # print 66 # bunzip2('repdata-data-StormData.csv.bz2') # require R.utils
}
```

Let's load the data into R. Then I added column *year* for analysing which years we have enough data for.
```{r}
library(R.utils)
library(knitr)

if (!("storm" %in% ls())) { # it will reduce rmd compilation time
  storm <- read.csv("d.csv") 
  #storm <- read.csv("repdata-data-StormData.csv")
} 
storm$year <- format(as.Date(storm$BGN_DATE, format = "%m/%d/%Y %H:%M:%S"), "%Y")
# View(table(storm$year))
#print(table(storm$year))


  kable(table(storm$year), format="markdown")


hist(as.numeric(storm[storm$year>1990 & storm$year<2011,]$year),
  xlab="Year", ylab="Number of events", main="Amount of data available by years")
```

Then I added column *year* for analysing which years we have enough data for:
```{r} 
#storm$year <- format(as.Date(storm$BGN_DATE, format = "%m/%d/%Y %H:%M:%S"), "%Y")
# View(table(storm$year))
```

Then we prepare a subset of data with only those variables that we will use for the study:
```{r}
d <- storm[storm$year>1990 & storm$year<2011,c("BGN_DATE", "EVTYPE","FATALITIES","INJURIES","PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP","year")]
# View(d[1:20,])
# View(table(storm$PROPDMGEXP))
hist(as.numeric(d$year))

#   write.table(d, "d.csv", sep=",", row.name=TRUE)
#   d <- read.csv("d.csv") 
#   View(head(d))
```

Aggregate data for injuries and fatalities:
```{r}
library(plyr)
fatalities <- ddply(d,~EVTYPE,summarise,t=sum(FATALITIES))
injuries <- ddply(d,~EVTYPE,summarise,t=sum(INJURIES))
fatalities <- tail(fatalities[order(fatalities$t),],10)
injuries <- tail(injuries[order(injuries$t),],10)
#View (fatalities)
```

### Most harmful to population health

You can see the types of events that appeared to be most harmful to population health:
```{r, echo=TRUE}
# library(knitr)
  kable(fatalities, format="markdown")
```

### types of events have the greatest economic consequences


The I calculated actual dollar amount for different types of damages. For this purpose I prepare manage multipliers: 
```{r}
table(d$PROPDMGEXP)
d$PROPDMGEXP <- sub("[-+?123456789H]","0", d$PROPDMGEXP, ignore.case=TRUE) # do not include "0" for replacement!
d$PROPDMGEXP <- sub("^$",  0, d$PROPDMGEXP, ignore.case=TRUE)
d$PROPDMGEXP <- sub("^B$", 9, d$PROPDMGEXP, ignore.case=TRUE)
d$PROPDMGEXP <- sub("^M$", 6, d$PROPDMGEXP, ignore.case=TRUE)
d$PROPDMGEXP <- sub("^K$", 3, d$PROPDMGEXP, ignore.case=TRUE)
#sum(d$PROPDMG)
#sum(d$PROPDMG * 10 ^ as.numeric(d$PROPDMGEXP))
d$PROPDMG <- d$PROPDMG * 10 ^ as.numeric(d$PROPDMGEXP)

table(d$CROPDMGEXP)
d$CROPDMGEXP <- sub("[-+?123456789H]","0", d$CROPDMGEXP, ignore.case=TRUE) # do not include "0" for replacement!
d$CROPDMGEXP <- sub("^$",  0, d$CROPDMGEXP, ignore.case=TRUE)
d$CROPDMGEXP <- sub("^B$", 9, d$CROPDMGEXP, ignore.case=TRUE)
d$CROPDMGEXP <- sub("^M$", 6, d$CROPDMGEXP, ignore.case=TRUE)
d$CROPDMGEXP <- sub("^K$", 3, d$CROPDMGEXP, ignore.case=TRUE)
d$CROPDMG <- d$CROPDMG * 10 ^ as.numeric(d$CROPDMGEXP)

dam1 <- ddply(d,~EVTYPE,summarise,t=sum(PROPDMG)) ## Property damage
dam2 <- ddply(d,~EVTYPE,summarise,t=sum(CROPDMG)) ## Agricultural (crop) damage

dam1 <- tail(dam1[order(dam1$t),],10) # prop
dam2 <- tail(dam2[order(dam2$t),],10) # crop
#View (dam1)


```

 

#Data

[https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2](Storm Data [47Mb])
There is also some documentation of the database available. Here you will find how some of the variables are constructed/defined.

National Weather Service Storm Data Documentation

National Climatic Data Center Storm Events FAQ

The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.
 
   


# Results

There should be a section titled Results in which your results are presented.
 
        
