---
title: "Tooth growth statistical inference"
author: "Stanislav Prikhodko"
date: "Friday, May 22, 2015"
output: html_document
---

## Loading data

Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r}
library(datasets)
library(ggplot2)
data(ToothGrowth) 
```
 
## Basic summary of the data

Here is a fragment of data that this study is built on:
```{r} 
head(ToothGrowth)
```

Here is some summary of this dataset:
```{r} 
summary(ToothGrowth)
```

In this studdy I will analyze the varriables *supp* and *dose* that have following values:
```{r} 
table(ToothGrowth$supp, ToothGrowth$dose)
```
 
## Exploratory data analysis

Looking at values of tooth length for different factors we can notice clear trends.

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
ggplot(ToothGrowth,aes(x=factor(dose),y=len,fill=dose)) + 
  geom_boxplot() +
  theme_bw() + 
  facet_wrap(~supp) +  
  scale_x_discrete("Dose [mg]") + 
  scale_y_continuous("Tooth length")
```

## Assumptions 

Since we have that low number of cases in the sample I will use **T-distribution** for the hypothesis testing.

## Confidence intervals and Hypothesis testing

Now, let's use confidence intervals to compare tooth growth by supp and dose. For this research I will prepare subsets of data by different doses:

```{r}
ToothDose1 <- subset(ToothGrowth, dose == 0.5)
ToothDose2 <- subset(ToothGrowth, dose == 1.0)
ToothDose3 <- subset(ToothGrowth, dose == 2.0)
```

Below is t-test for the variable **supp** within all doses and in each of subsets by different doses:
```{r}
t.test(len ~ supp, data = ToothGrowth)
```

```{r}
t.test(len ~ supp, data = ToothDose1) ## dose = 0.5
```

```{r}
t.test(len ~ supp, data = ToothDose2) ## dose = 1.0
```

```{r}
t.test(len ~ supp, data = ToothDose3) ## dose = 2.0
```

## Conclusions 

Looking at all groups together and the group of samples with dose=2.0 we see that p-value is greater than 0.05 (and the confidence interval contains 0) so we cannot rejecct the null hypothesis. 

However, in the group of samples where dose = 0.5 and 1.0 we are 95% confident that there is a difference between true mean of *tooth length* in groups OJ and VC. 

