---
title: "Tooth growth statistical inference"
author: "Stanislav Prikhodko"
date: "Friday, May 22, 2015"
output: pdf_document
---

## Loading data

Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r}
library(datasets)
library(ggplot2)
data(ToothGrowth) 
```
 
## Basic summary of the data

Here is a fragment of data that this study is built on:
```{r} 
head(ToothGrowth)
```

Here is some summary of this dataset:
```{r} 
summary(ToothGrowth)
```

In this studdy I will analyze the varriables *supp* and *dose* that have following values:
```{r} 
table(ToothGrowth$supp, ToothGrowth$dose)
```
 
## Exploratory data analysis

Looking at values of tooth length for different factors we can notice clear trends.

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
ggplot(ToothGrowth,aes(x=factor(dose),y=len,fill=dose)) + 
  geom_boxplot() +
  theme_bw() + 
  facet_wrap(~supp) +  
  scale_x_discrete("Dose [mg]") + 
  scale_y_continuous("Tooth length")
```

## Assumptions 

Since we have that low number of cases in the sample I will use **T-distribution** for the hypothesis testing.

## Confidence intervals and Hypothesis testing

Now, let's use confidence intervals to compare tooth growth by supp and dose. For this research I will prepare subsets of data by different doses:

```{r}
ToothDose1 <- subset(ToothGrowth, dose == 0.5)
ToothDose2 <- subset(ToothGrowth, dose == 1.0)
ToothDose3 <- subset(ToothGrowth, dose == 2.0)
```

Below is t-test for the variable **supp** within all doses:
```{r}
t.test(len ~ supp, data = ToothGrowth)
```

## Conclusions 

State your conclusions and the assumptions needed for your conclusions. 
Were the results of the tests and/or intervals interpreted in the context of the problem correctly?

 

